\chapter*{Foreword}  

The TCCM master's programme - and this thesis project in particular - provided me with the opportunity to open a completely new chapter of my life. Not only did it allow me to study at a top-notch university, but, more importantly, it gave me the chance to be surrounded by brilliant people who inspired me to think outside the box and to learn many new things, not merely in a professional sense, but also on a personal level.

\par\smallskip
I'm wholeheartedly grateful to my supervisor, Prof. Jeremy Harvey, who agreed to take me on from the time of my internship in the first year and continued to support me throughout my master's thesis. His constant guidance and encouragement made this project possible. However, most significantly, our discussions and brainstorming sessions had the greatest impact on me. They helped me to think critically about my research topic and to consider it from different perspectives.

\par\smallskip
I'd like to thank my family in Uzbekistan for always being there for me and supporting me throughout my studies. Their love and support helped me get through all the ups and downs of living in a completely new environment. \begin{otherlanguage}{russian} Большое спасибо моей семье: маме, бабушке, папе и младшему брату за веру в меня. Люблю вас всех и крепко обнимаю! Без вашей поддержки у меня бы точно не получилось начать новую жизнь в совершенно другой части мира. \end{otherlanguage}

\par\smallskip
Student life is not only about studying, but also about making friends and having fun. In this regard, I would like to thank those I hold dear: my friends, all group and division members, as well as the TCCM cohort, for making these two years an unforgettable experience. Special \textit{dank u wel} and \textit{muchas gracias} go to everyone for the coffee breaks we shared - they were not only a great way to unwind, but also a source of new ideas.

\par\smallskip
Last but not least, I would like to express my gratitude to my former supervisors and colleagues in Uzbekistan, who motivated me to embark on an academic journey and to pursue a master's degree abroad. For that, I would like to thank Dr. Artyom Baev, Dr. Anvar Sariev, Prof. Zukhra Kadirova, and Prof. Shahlo Daminova. 

\par\smallskip
This project would not have been possible without the computational resources provided by KU Leuven and the Erasmus+ scholarship from the EMJMD TCCM.


\chapter*{Contribution statement}
Albert Makhmudov proposed revisiting phosphate hydrolysis using a more advanced computational setup. Prof. Jeremy Harvey suggested the investigation of a system containing methyl diphosphate and proposed the use of a neural network potential to describe the reaction mechanism. The overall workflow was designed together by Albert Makhmudov and Prof. Jeremy Harvey, who also jointly analysed the results. All calculations, figures, and visualisations were carried out by Albert Makhmudov. The thesis was written by Albert Makhmudov with feedback and corrections from Prof. Jeremy Harvey. Ehsan Moravveji (VSC, KU Leuven), Hans Vansweevelt (KU Leuven), and Anders Johansson (Harvard) helped with the installing and compiling the software on the high-performance computing clusters.